\textsc{\wobble: a data-driven method for precision radial velocities}

### 2018-12-04 paper/paper.tex

For the purposes of this work, we chose to test the performance of \wobble on archival spectra from the High Accuracy Radial Velocity Planet Searcher (\HARPS) spectrograph \citep{Mayor2003}. 
\HARPS has operated continuously since 2003 and as such has an extensive catalog of publicly available data. 
Furthermore, its excellent instrumental stability and careful calibration make the data ideally suited to our method, which relies on having an accurate wavelength solution for every spectrum.

All data were obtained from the \acronym{ESO} public data archive in the form of ``\code{e2ds}'' spectra. 
These data come as extracted 1D spectra, in blocks of 72 orders by 4096 pixels per order.
The airmass of the observation and the calculated \BERV are both provided in the \acronym{FITS} header of the spectral data file. 
We obtained wavelength solutions for each spectrum from the dedicated \HARPS calibration archive, also maintained by \acronym{ESO}. 
\todo{Uncertainty estimates for each pixel were calculated as... [HOGG, help me fill this in - relation to \SNR and conversion to log]}

Before running the above-described model optimization on the data, we first mask out any unreliably measured spectral regions and do a continuum normalization on each spectrum. 
This masking is done by setting the inverse variance on the masked data point to zero, effectively removing it from the fit. 
In brief, we mask pixels whose extracted flux is below zero as well as regions at the edges of the spectral orders where the local \SNR falls below 5. 
After masking these data, we convert the flux to log space and continuum normalize by fitting a polynomial to an asymmetrically clipped subset of the data.

@@

In general, the data used in this work are high quality in terms of \SNR, spectral resolution, and quantity. 
As demonstrated for the case of \Mdwarf, \wobble is limited in its operability when the data are at low \SNR. 
\wobble implicitly assumes Gaussian noise in logarithmic flux, which is not strictly accurate. 
This may become a substantial issue at low \SNR, making \wobble a poor tool choice for accurate modeling of very low \SNR spectra. 
\todo{(HOGG: maybe say something about how we could mitigate this with changes to continuum handling?)}

Our method may also fail at low spectral resolution. 
In particular, if the stellar spectrum does not shift with respect to the telluric spectrum by at least a resolution element, it will become extremely challenging to disentangle the two components. 
\todo{(HOGG: say other things about undersampled lines?)}

@@

It is likely that for such an instrument, our assumption that the wavelength solution is perfect does not hold. 
One simple way to deal with this would be to allow the observatory rest frame component(s) of the model to change in \RV. 
The stellar \RV measurement would then be differential with respect to the effective spectrograph \RV. 
This effective spectrograph \RV may vary in different spectral orders or smaller wavelength regions. \todo{(HOGG: say a thing)} 
\todo{(TO DO: add some citations to using tellurics for this)}

>>>>

For the purposes of this work, we chose to test the performance of \wobble on archival spectra from the High Accuracy Radial Velocity Planet Searcher (\HARPS) spectrograph \citep{Mayor2003}. 
\HARPS has operated continuously since 2003 and as such has an extensive catalog of publicly available data. 
Furthermore, its excellent instrumental stability and precise calibration make the data ideally suited to our method, which relies on having an accurate wavelength solution for every spectrum.

All data were obtained from the \acronym{ESO} public data archive in the form of ``\code{e2ds}'' spectra. 
These data come as extracted 1D spectra, in blocks of 72 orders by 4096 pixels per order.
The airmass of the observation and the calculated \BERV are both provided in the \acronym{FITS} header of the spectral data file. 
We obtained wavelength solutions for each spectrum from the dedicated \HARPS calibration archive, also maintained by \acronym{ESO}.

Before running the above-described model optimization on the data, we first mask out any unreliably measured spectral regions and do a continuum normalization on each spectrum. 
This masking is done by setting the inverse variance on the masked data point to zero, effectively removing it from the fit. 
In brief, we mask pixels whose extracted flux is below zero as well as regions at the edges of the spectral orders where the local \SNR falls below 5. 
After masking these data, we convert the flux to log space and continuum normalize by fitting and subtracting a polynomial to an asymmetrically clipped subset of the data.

To each pixel in the normalized, logarithmic spectrum, we assign an uncertainty variance (squared error).
Technically this uncertainty is assumed to be Gaussian in the log space, which is an incorrect approximation, but not very far off for high signal-to-noise data.
\todo{BEDELL: Check this:}
For each order we assign to each log flux an uncertainty variance that is the inverse of the mean signal-to-noise-squared of the pixels in that order (as reported by the \HARPS pipelines), scaled up or down by the raw linear flux observed in that pixel in the data prior to logging and normalizing.
This is the best Poisson estimate we can make given that we are working in the log space, and we don't have individual-pixel uncertainty estimates.
Technically a small bias is introduced by taking the logarithm of the flux, but this bias is very small at the relevant signal-to-noise ratios; furthermore, the small bias acts in the flux direction, not the radial-velocity direction.

@@

In general, the data used in this work are high quality in terms of \SNR, spectral resolution, and quantity. 
As demonstrated for the case of \Mdwarf, \wobble is limited in its operability when the data are low in \SNR. 
\wobble implicitly assumes Gaussian noise in logarithmic flux, which is not strictly accurate. 
This could become a substantial issue at low \SNR, making \wobble a poor tool choice for accurate modeling of very low \SNR spectra.
At this point, it is also the case that our continuum-normalization method also does poorly at low \SNR, so there are various reasons \wobble is not optimal for faint sources or very short exposure times.
However, the fact that the model is successful in the form of a linear additive model in the log space means that it could probably be converted to work as a multiplicative model in the linear space; that conversion and investigation is beyond the scope of the current work, which is optimized for performance on typical \HARPS observing campaigns.

Our method may also fail at low spectral resolution. 
In particular, if the stellar spectrum does not shift with respect to the telluric spectrum by at least a resolution element (naively this is expected to become a problem at resolutions below $10^4$), it could become challenging to disentangle the two components.
That said, there still is a causal difference in the data space between signals that are fixed with respect to the star and signals that are fixed with respect to the atmosphere; in principle it is a question not purely of resolution but a combination of resolution and \SNR.
Naively, at low resolution, the model is expected to work well when the total signal-to-noise in the data set (greatly) exceeds the dimensionless ratio of the line width to the radial-velocity variation.

@@

It is likely that for such an instrument, our assumption that the wavelength solution is perfect does not hold; hence the cell. 
One simple way to deal with this would be to allow the observatory rest frame component(s) of the model to change in \RV. 
The stellar \RV measurement would then be differential with respect to the effective spectrograph \RV. 
This effective spectrograph \RV may vary in different spectral orders or smaller wavelength regions.
The most extreme extension would be to include the full wavelength solution as model parameters, and optimize them along with the spectral components and radial velocities.
\todo{(TO DO: add some citations to using tellurics for this)}
